\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2023-07-02 cubble.Rmd
make it clearer the attributes apply to both cubble subclasses

A cubble object inherits the attributes from tibble (and its subclasses): \code{class}, \code{row.names}, and \code{names}. Additionally, it has three specialised attributes: \code{key}, \code{index}, and \code{coords}. \code{key} and \code{index} are used as introduced in the \pkg{tsibble} package [@tsibble]. In cubble, the \code{key} attribute identifies the row in the spatial cubble (given the internal use of \code{tidyr::nest()} for nesting), and when combined with the \code{index} argument, it identifies the row in the temporal cubble. Currently, cubble only supports one variable as the key, and the accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes. The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value. The \pkg{sf} package [@sf] is used under the hood to calculate the spatial bounding box, displayed in the header of a spatial cubble.

The temporal cubble has a special attribute called \code{spatial} to store the spatial variables. Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:

>>>>

Both cubble objects inherit the attributes \code{class}, \code{row.names}, and \code{names} from tibble, which originates from data frames. Additionally, both have three specialised attributes: \code{key}, \code{index}, and \code{coords}. \code{key} and \code{index} are used as introduced in the \pkg{tsibble} package [@tsibble]. In cubble, the \code{key} attribute identifies the row in the spatial cubble (given the internal use of \code{tidyr::nest()} for nesting), and when combined with the \code{index} argument, it identifies the row in the temporal cubble. Currently, cubble only supports one variable as the key, and the accepted temporal classes for \code{index} includes the base R classes \code{Date}, \code{POSIXlt}, \code{POSIXct}, as well as tsibble's \code{yearmonth}, \code{yearweek}, and \code{yearquarter} classes. The \code{coords} attribute represents an ordered pair of coordinates that can be either an unprojected pair of longitude and latitude, or a projected easting and northing value. Moreover, temporal cubbles have a special attribute called \code{spatial} to store the spatial variables. Shortcut functions are available to extract attributes from the temporal cubble object, for example, \code{spatial()} for extracting spatial variables:
